---
title: "Compare Approaches"
author: "Melissa Wong"
date: "12/15/2021"
output: html_document
---

# Problem Description

Given some observational data and the following assumed generative model, we are interested in estimating the effect of _x_ (i.e., $\beta$).

\begin{align*}
  y &\sim N(\mu, \sigma) \\
  \mu &= \alpha + x*\beta
\end{align*}

However, after doing some exploratory data analysis it appears that some of the response values are truncated or rounded (we don't know which). The question is how does the proportion of truncated/rounded data affect the estimate of $\beta$ and at what point can we no longer ignore the truncation/rounding?

Let's consider the following approaches:

1. Kolmogorov-Smirnov test

2. Shapiro-Wilk test

3. OLS Confidence Interval Coverage

4. Bayesian Credible Interval Coverage

# Setup

```{r libraries, results='hide', message=FALSE, warning=FALSE}

library(tidyverse)
library(broom)
library(RColorBrewer)
library(here)
```

```{r options}
options("scipen" = 1, "digits" = 4)

#knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(message = FALSE)
#knitr::opts_chunk$set(warning = FALSE)
#knitr::opts_chunk$set(out.width = "50%")
#knitr::opts_chunk$set(fig.align = "center")

set.seed(1234)
```

# User-defined Parameters

```{r user_params}
# Proportion of rounded/truncated data
prop <- seq(0.1, 0.9, 0.1)

# Number of obs
N <- c(25, 100, 500, 1000)

params <- expand.grid(N=N, prop=prop)

# True parameter values
alpha_true = -1
beta_true = 1
sigma_true = 1
```

# Define Helper Functions

```{r sim_data}

#' Simulate data
#'
#' This simulates true/reference data from a Normal Distribution
#' and then truncated and rounded subsets based on parameters passed in.
#'
#' @return data.frame containing y_true, y_trunc and y_round.
#' 
sim_data <- function(N, p, a, b, sigma)
{
    # Simulate data
    tmp <- data.frame(x = rnorm(N, 0, 5)) %>% 
      mutate(
        y_true = a + b*x + rnorm(N, 0, sigma),
         y_round = y_true,
         y_trunc = y_true
      )
    
    sidx <- sample(1:N, N * p, replace=FALSE)
    
    # Simulate a "mixture" with some subset of the data truncated/rounded
    tmp$y_round[sidx] <- round(tmp$y_round[sidx])
    tmp$y_trunc[sidx] <- trunc(tmp$y_trunc[sidx])
    
    return(tmp)
}
```


```{r ci_contains}

#' Check if CI contains true/reference values
#'
#' Given a tidy data.frame, check if the (1-alpha)% CI contains the true value.
#'
#' @return boolean.
#' 
ci_contains <- function(fit, true, alpha=0.05) {
  lb <- fit$estimate + qnorm(alpha/2) * fit$std.error
  ub <- fit$estimate - qnorm(alpha/2) * fit$std.error
  return((lb <= true) & (true <= ub))
}
```


```{r reject_rates}

#' Runs simulations to estimate rejection rates for various methods
#'
#' 
#'
#' 
reject_rates <- function(N, p, a, b, sigma, alpha=0.05)
{
  Nsim <- 1000
  
  KS_round_reject <- rep(NA, Nsim)
  KS_trunc_reject <- rep(NA, Nsim)
  SW_round_reject <- rep(NA, Nsim)
  SW_trunc_reject <- rep(NA, Nsim)
  # OLS_round_reject <- rep(NA, Nsim)
  # OLS_trunc_reject <- rep(NA, Nsim)
  
  for (i in 1:Nsim) {

    tmp <- sim_data(N, p, a, b, sigma)
    
    
    # Kolmogorov-Smirnov Tests
    KS_round_reject[i] <- tidy(ks.test(tmp$y_round, "pnorm", 
                                    mean=mean(tmp$y_round), 
                                    sd=sd(tmp$y_round))
                            )$p.value < alpha
    
    KS_trunc_reject[i] <- tidy(ks.test(tmp$y_trunc, "pnorm", 
                                    mean=mean(tmp$y_trunc), 
                                    sd=sd(tmp$y_trunc))
                            )$p.value < alpha
    
    # Shapiro-Wilk Tests
    SW_round_reject[i] <- tidy(shapiro.test(tmp$y_round))$p.value < alpha
    
    SW_trunc_reject[i] <- tidy(shapiro.test(tmp$y_trunc))$p.value < alpha
    
    # OLS CIs
    
    # fit.round <- tidy(lm(y_round ~ x, data=tmp)) %>% 
    #   filter(term == "x")
    # OLS_round_reject[i] <- ci_contains(fit.round, b)
    # 
    # fit.trunc <- tidy(lm(y_trunc ~ x, data=tmp)) %>% 
    #   filter(term == "x")
    # 
    # OLS_trunc_reject[i] <- ci_contains(fit.trunc, b)
  }
  
  return(list(KS_round_reject = mean(KS_round_reject),
         KS_trunc_reject = mean(KS_trunc_reject),
         SW_round_reject = mean(SW_round_reject),
         SW_trunc_reject = mean(SW_trunc_reject))
  )
}

```



```{r coverage_rates}

#' Runs simulations to estimate rejection rates for various methods
#'
#' 
#'
#' 
coverage_rates <- function(N, p, a, b, sigma, alpha=0.05)
{
  Nsim <- 1000
  
  OLS_round_reject <- rep(NA, Nsim)
  OLS_trunc_reject <- rep(NA, Nsim)
  
  for (i in 1:Nsim) {

    tmp <- sim_data(N, p, a, b, sigma)
    
    # OLS CIs
    
    fit.round <- tidy(lm(y_round ~ x, data=tmp)) %>%
      filter(term == "x")
    OLS_round_reject[i] <- ci_contains(fit.round, b)

    fit.trunc <- tidy(lm(y_trunc ~ x, data=tmp)) %>%
      filter(term == "x")

    OLS_trunc_reject[i] <- ci_contains(fit.trunc, b)
  }
  
  return(list(OLS_round_reject = mean(OLS_round_reject),
         OLS_trunc_reject = mean(OLS_trunc_reject))
  )
}

```

# Tests for Normality

```{r KS_test, warning=FALSE}

res <- apply(params, 1, function(row) reject_rates(row[1], 
                                                row[2],
                                                alpha_true,
                                                beta_true,
                                                sigma_true)) 

cbind(params,
      do.call(rbind.data.frame, res)) %>% 
  mutate(N=as.factor(N)) %>% 
  pivot_longer(-c("N", "prop"), names_to="type") %>% 
  separate(type, c("test_type", "data_type", NULL)) %>% 
  ggplot(mapping=aes(x=prop, 
                     y=value,
                     group=N,
                     color=N)) +
  geom_point() +
  geom_line() +
  facet_grid(test_type ~ data_type) +
  labs(x="Proportion of data rounded/truncated",
       y="Prob[reject Ho]",
       color="Number of obs",
       title="Rejection rates, One-sample KS-Test vs SW-Test") +
  scale_x_continuous(breaks=prop)
```

# Actual Coverage of OLS Confidence Intervals

```{r}
res <- apply(params, 1, function(row) coverage_rates(row[1], 
                                                row[2],
                                                alpha_true,
                                                beta_true,
                                                sigma_true)) 

cbind(params,
      do.call(rbind.data.frame, res)) %>% 
  mutate(N=as.factor(N)) %>% 
  pivot_longer(-c("N", "prop"), names_to="type") %>% 
  separate(type, c("test_type", "data_type", NULL)) %>% 
  ggplot(mapping=aes(x=prop, 
                     y=value,
                     group=N,
                     color=N)) +
  geom_point() +
  geom_line() +
  geom_hline(mapping=aes(yintercept=0.95),
             color="red",
             linetype="dashed") +
  facet_grid(test_type ~ data_type) +
  labs(x="Proportion of data rounded/truncated",
       y="Proportion CIs Containg True Value",
       color="Number of obs",
       title="CI Actual Coverage") +
  scale_x_continuous(breaks=prop)
```

